---
title: 'Lab 4: Data Frame Basics'
subtitle: "Stat 133, Fall 2017"
author: "Sudarshan Srirangapatanam"
output: github_document
---

## Creating data frames
```{R create}
require("knitr")

player_name <- c("Thompson", "Curry", "Green", "Durant", "Pachulia")
pos <- c("SG", "PG", "PF", "SF", "C")
sal <- c(16663575, 12112359, 15330435, 26540100, 2898000)
point <- c(1742, 1999, 776, 1555, 426)
PPG <- c(22.3, 25.3, 10.2, 25.1, 6.1)
rookie <- c(FALSE, FALSE, FALSE, FALSE, FALSE)

lineup_GSW <- data.frame(
  Player = player_name,
  Position = pos,
  Salary = sal,
  Points = point,
  PPG = PPG,
  Rookie = rookie
)

gsw_list <- list(
  Player = player_name,
  Position = pos,
  Salary = sal,
  Points = point,
  PPG = PPG,
  Rookie = rookie
)

lineup_GSW_list <- data.frame(gsw_list)

lineup_cbind <- cbind(
    Player = player_name,
  Position = pos,
  Salary = sal,
  Points = point,
  PPG = PPG,
  Rookie = rookie
)
```

+ Use the vectors to create a first data frame with `data.frame()`.<br>
`r kable(lineup_GSW)`

+ Create another data frame by first starting with a `list()`, and then passing the list to `data.frame()`.<br>
`r kable(lineup_GSW_list)`

+ What would you do to obtain a data frame such that when you check its structure `str()` the variables are:
    + Player as character: enclose our vector when we pass them with `as.character`.
    + Position as factor: enclose our vector when we pass them with `as.factor`.
    + Salary as numeric or real (ignore the commas): enclose our vector when we pass them with `as.numeric`, or `as.real`.
    + Points as integer: enclose our vector when we pass them with `as.integer`.
    + PPG as numeric or real: enclose our vector when we pass them with `as.numeric`, or `as.real`.
    + Rookie as logical: enclose our vector when we pass them with `as.logical`.

+ Find out how to use the column binding function `cbind()` to create a tabular object with the vectors created in step 1.<br>
`r kable(lineup_cbind)`
    + inspect what class of object is obtained with `cbind()`: `r class(lineup_cbind)`
    
+ How could you convert the object in the previous step into a data frame?<br>
We could coerce it using `as.data.frame()` function.

## Data
```{r data_read}
dat <- read.csv('data/nba2017-players.csv', stringsAsFactors = FALSE)

durant <- dat[dat$player == "Kevin Durant",]

ucla <- dat[dat$college == "University of California, Los Angeles",]

rookies <- dat[dat$experience == 0,]
rookie_centers <- rookies[rookies$position == "C",]

top_players <- dat[dat$games > 50 & dat$minutes > 100,]

gsw <- dat[dat$team == "GSW", c("player", "height", "weight")]
```

**The data used in this lab is taken from [nba2017-players](https://github.com/Sudarshan-UCB/stat-133-fall2017/blob/master/labs/Lab04/data/nba2017-players.csv)**

+ `dim()`:<br>
`r dim(dat)`
  
+ `head()`:<br>
`r kable(head(dat), row.names = FALSE)`
  
+ `tail()`:<br>
`r kable(tail(dat), row.names = FALSE)`
  
+ `summary()`:<br>
`r kable(summary(dat), row.names = FALSE)`

Use bracket notation, the dollar operator, as well as concepts of logical subsetting and indexing to:

+ Display the last 5 rows of the data.<br>
`r kable(dat[(length(dat)-5+1):length(dat),], row.names = FALSE)`

+ Display those rows associated to players having height less than 70 inches tall.<br>
`r kable(dat[dat$height < 70,], row.names = FALSE)`

+ Of those players that are centers (position C), display their names and salaries.\*<br>
`r kable(head(dat[dat$position == "C",c("player", "salary")]), row.names = FALSE)`

+ Create a data frame durant with Kevin Durant's information (i.e. row).<br>
`r kable(durant, row.names = FALSE)`

+ Create a data frame ucla with the data of players from college UCLA ("University of California, Los Angeles").\*<br>
`r kable(head(ucla), row.names = FALSE)`

+ Create a data frame rookies with those players with 0 years of experience.\*<br>
`r kable(head(rookies), row.names = FALSE)`

+ Create a data frame rookie_centers with the data of Center rookie players.\*<br>
`r kable(head(rookie_centers), row.names = FALSE)`

+ Create a data frame top_players for players with more than 50 games and more than 100 minutes played.\*<br>
`r kable(head(top_players), row.names = FALSE)`

+ What's the largest height value?<br>
`r max(dat$height)`

+ What's the minimum height value?<br>
`r min(dat$height)`

+ What's the overall average height?<br>
`r mean(dat$height)`

+ Who is the tallest player?<br>
`r dat[dat$height == max(dat$height), "player"]`

+ Who is the shortest player?<br>
`r dat[dat$height == min(dat$height), "player"]`

+ Which are the unique teams?<br>
`r unique(dat$team)`

+ How many different teams?<br>
`r length(unique(dat$team))`

+ Who is the oldest player?<br>
`r dat[dat$age == max(dat$age), "player"]`

+ What is the median salary of all players?<br>
`r median(dat$salary)`

+ What is the median salary of the players with 10 years of experience or more?<br>
`r median(dat[dat$experience >= 10,"salary"])`

+ What is the median salary of Shooting Guards (SG) and Point Guards (PG)?<br>
`r median(dat[dat$position == "SG" | dat$position == "PG","salary"])`

+ What is the median salary of Power Forwards (PF), 29 years or older, and 74 inches tall or less?<br>
`r median(dat[dat$position == "PF" & dat$age >= 29 & dat$height <= 74,"salary"])`

+ How many players scored 4 points or less?<br>
`r length(dat[dat$points <= 4,"player"])`

+ Who are those players who scored 4 points or less?<br>
`r dat[dat$points <= 4,"player"]`

+ Who is the player with 0 points?<br>
`r dat[dat$points == 0,"player"]`

+ How many players are from "University of California, Berkeley"?<br>
`r length(dat[dat$college == "University of California, Berkeley","player"])`

+ Are there any players from "University of Notre Dame"? If so how many and who are they?<br>
There are `r length(dat[dat$college == "University of Notre Dame","player"])` players from University of Notre Dame. They are `r dat[dat$college == "University of Notre Dame","player"]`.

+ Are there any players with weight greater than 260 pounds? If so how many and who are they?<br>
There are `r length(dat[dat$weight > 260,"player"])` players with weight greater than 260 pounds.

+ How many players did not attend a college in the US?<br>
`r length(dat[dat$college == "","player"])`

+ Who is the player with the maximum rate of points per minute?<br>
`r transform(dat, PPM = points/minutes)[transform(dat, PPM = points/minutes)$PPM == max(transform(dat, PPM = points/minutes)$PPM), "player"]`

+ Who is the player with the maximum rate of three-points per minute?<br>
`r transform(dat, PPM = points3/minutes)[transform(dat, PPM = points3/minutes)$PPM == max(transform(dat, PPM = points3/minutes)$PPM), "player"]`

+ Who is the player with the maximum rate of two-points per minute?<br>
`r transform(dat, PPM = points2/minutes)[transform(dat, PPM = points2/minutes)$PPM == max(transform(dat, PPM = points2/minutes)$PPM), "player"]`

+ Who is the player with the maximum rate of one-points (free-throws) per minute?<br>
`r transform(dat, PPM = points1/minutes)[transform(dat, PPM = points1/minutes)$PPM == max(transform(dat, PPM = points1/minutes)$PPM), "player"]`

+ Create a data frame gsw with the name, height, weight of Golden State Warriors (GSW)\*<br>
`r kable(head(gsw), row.names = FALSE)`

+ Display the data in gsw sorted by height in increasing order (hint: see ?sort and ?order)\*<br>
`r kable(head(gsw[order(gsw$height),]), row.names = FALSE)`

+ Display the data in gsw by weight in decreasing order (hint: see ?sort and ?order)\*<br>
`r kable(head(gsw[order(-gsw$weight),]), row.names = FALSE)`

+ Display the player name, team, and salary, of the top 5 highest-paid players (hint: see ?sort and ?order)<br>
`r kable(dat[order(-dat$salary),c("player", "team", "salary")][1:5,], row.names = FALSE)`

+ Display the player name, team, and points3, of the top 10 three-point players (hint: see ?sort and ?order)<br>
`r kable(dat[order(-dat$points3),c("player", "team", "points3")][1:10,], row.names = FALSE)`

*\* only a small suset is shown using the `head()` function to reduce report length*

## Group by
```{r group_by}
pos_group <- aggregate(. ~ position, data = dat[ ,c('position', 'height', 'weight', 'age')], FUN = mean)

team_group <- aggregate(. ~ team, data = dat[ ,c('team', 'height', 'weight', 'age')], FUN = mean)

team_pos_group <- aggregate(. ~ team + position, data = dat[ ,c('team', 'position', 'height', 'weight', 'age')], FUN = mean)

fun_group <- aggregate(. ~ team + position, data = dat[ ,c('team', 'position', 'salary')], FUN = function(x) c(minimum = min(x), median = median(x), mean = mean(x), maximum = max(x)))
```

+ Create a data frame with the average height, average weight, and average age, grouped by position\*<br>
`r kable(head(pos_group), row.names = FALSE)`

+ Create a data frame with the average height, average weight, and average age, grouped by team\*<br>
`r kable(head(team_group), row.names = FALSE)`

+ Create a data frame with the average height, average weight, and average age, grouped by team and position.\*<br>
`r kable(head(team_pos_group), row.names = FALSE)`

+ Difficult: Create a data frame with the minimum salary, median salary, mean salary, and maximum salary, grouped by team and position.\*<br>
```{r}
head(fun_group)
```

*\* only a small suset is shown using the `head()` function to reduce report length*


