---
title: 'Lab 7: Simple Functions and Conditionals'
subtitle: "Stat 133, Fall 2017"
author: "Sudarshan Srirangapatanam"
output: github_document
---

## Setup

```{r setup}
library("knitr")
library("dplyr")
library("ggplot2")
library("readr")

knitr::opts_chunk$set(fig.path = "../images/")
```

+ `knitr` is used for knitting the document as well as other fucntions to cleanup the output
+ `readr` is used for reading data into R
+ `dplyr` is used for data wrangling
+ `ggplot2` is used to generate any necessary plots

```{r fxns}
f <- function(x) {
  x^2
}

g <- function(x){
  (2*x) +5
}

fog <- function(x){
  g_x <- g(x)
  f(g_x)
}

gof <- function(x){
  f_x <- f(x)
  g(f_x)
}

pythagoras <- function(a, b = a){
  sqrt(a^2 + b^2)
}

miles2kms <- function(a = 1){
  a * 1.6
}

gallons2liters <- function(g = 1){
  g * 3.78541
}

liters2gallons <- function(l = 1){
  gallons2liters(l) / gallons2liters(1)^2
}

seconds2years <- function(sec){
  min <- (1/60) * sec
  hour <- (1/60) * min
  day <- (1/24) * hour
  year <- (1/365.25) * day
  year
}

gaussian <- function(x, m, s){
  exp(-.5 * ((x - m) / s)^2) / (s * sqrt(2*pi))
}
```

+ f(2): `r f(2)`
+ f(-5): `r f(-5)`

+ g(0): `r g(0)`
+ g(-5/2): `r g(-5/2)`

+ fog(2): `r fog(2)`
+ fog(-5): `r fog(-5)`

+ gof(0): `r gof(0)`
+ gof(-5/2): `r gof(-5/2)`

+ pythagoras(3, 4): `r pythagoras(3, 4)`
+ pythagoras(5): `r pythagoras(5)`

+ testing `miles2kms()`:
`r a <- c(1:10, seq(20,100,10)); kable(data.frame("miles" = a, "kms" = miles2kms(a)))`

+ testing `liters2gallons()`:
`r a <- c(1:10, seq(20,100,10)); kable(data.frame("liters" = a, "gallons" = liters2gallons(a)))`

+ seconds2years(1000000000) `r seconds2years(1000000000)`<br>
  Can a newborn baby in USA expect to live for one billion (10^9) seconds?<br>
  Yes, because life expectancy is 79 years.
  
+ `gaussian()` evaluated at x = 1, m = 0, and s = 2: `r gaussian(1, 0, 2)`<br>
`dnorm()` fucntion: `r dnorm(1, 0, 2)`

```{r gauss_plot}
x_values <- seq(-4.5, 4.5, .1)
y_values <- gaussian(x_values, 0, 2)
plot(x_values, y_values, type = "l", lwd = 2)
```

