---
title: 'Lab 7: Simple Functions and Conditionals'
subtitle: "Stat 133, Fall 2017"
author: "Sudarshan Srirangapatanam"
output: github_document
---

## Setup

```{r setup}
library("knitr")
library("dplyr")
library("ggplot2")
library("readr")

knitr::opts_chunk$set(fig.path = "../images/")
```

+ `knitr` is used for knitting the document as well as other fucntions to cleanup the output
+ `readr` is used for reading data into R
+ `dplyr` is used for data wrangling
+ `ggplot2` is used to generate any necessary plots

```{r}
f <- function(x){
  x^2
}

g <- function(x){
  (2*x) +5
}

fog <- function(x){
  g_x <- g(x)
  f(g_x)
}

gof <- function(x){
  f_x <- f(x)
  g(f_x)
}

pythagoras <- function(a, b = a){
  sqrt(a^2 + b^2)
}
```

+ f(2): `r f(2)`
+ f(-5): `r f(-5)`

+ g(0): `r g(0)`
+ g(-5/2): `r g(-5/2)`

+ fog(2): `r fog(2)`
+ fog(-5): `r fog(-5)`

+ gof(0): `r gof(0)`
+ gof(-5/2): `r gof(-5/2)`

+ pythagoras(3, 4): `r pythagoras(3, 4)`
+ pythagoras(5): `r pythagoras(5)`

```{r}
miles2kms <- function(a = 1){
  a * 1.6
}

gallons2liters <- function(g = 1){
  g * 3.78541
}

liters2gallons <- function(l = 1){
  gallons2liters(l) / gallons2liters(1)^2
}

seconds2years <- function(sec){
  min <- (1/60) * sec
  hour <- (1/60) * min
  day <- (1/24) * hour
  year <- (1/365.25) * day
  year
}
```

+ testing `miles2kms()`:
`r a <- c(1:10, seq(20,100,10)); kable(data.frame("miles" = a, "kms" = miles2kms(a)))`

+ testing `liters2gallons()`:
`r a <- c(1:10, seq(20,100,10)); kable(data.frame("liters" = a, "gallons" = liters2gallons(a)))`

+ seconds2years(1000000000) `r seconds2years(1000000000)`<br>
  Can a newborn baby in USA expect to live for one billion (10^9) seconds?<br>
  Yes, because life expectancy is 79 years.

```{r}
gaussian <- function(x, m, s){
  exp(-.5 * ((x - m) / s)^2) / (s * sqrt(2*pi))
}

plot_points <- function(fxn){
  x <- seq(-4, 4, length.out = 20)
  y <- fxn(x)
  plot(x, y, type = "l", lwd = 3, col = "red")
  abline(h = 0, v = 0)
  title(main = paste(body(fxn))[2])
}

poly1 <- function(x){
  (x^2) * (x - 1)
}

poly2 <- function(x){
  x^3
}

poly3 <- function(x){
  (x^2 - 1) * (x + 3)^3
}

poly4 <- function(x){
  (x^2 - 1) * (x^2 - 9)
}
```
  
+ `gaussian()` evaluated at x = 1, m = 0, and s = 2: `r gaussian(1, 0, 2)`<br>
`dnorm()` fucntion: `r dnorm(1, 0, 2)`

```{r}
x_values <- seq(-4.5, 4.5, .1)
y_values <- gaussian(x_values, 0, 2)
plot(x_values, y_values, type = "l", lwd = 2)

plot_points(poly1)
plot_points(poly2)
plot_points(poly3)
plot_points(poly4)
```

```{r}
descriptive <- function(x){
  out <- c("min" = min(x),
           "q1" = quantile(x, probs = .25),
           "median" = median(x),
           "mean" = mean(x),
           "q3" = quantile(x, probs = .75),
           "max" = max(x),
           "range" = range(x),
           "iqr" = IQR(x),
           "sd" = sd(x))
  return(out)
}

combinations <- function(n, k){
  factorial(n) / (factorial(k) * factorial(n - k))
}

binom_prob <- function(n, k, prob){
  combinations(n, k) * prob^k * (1 - prob)^(n-k)
}
```

+ testing `descriptive()`:<br>
  `r x <- rnorm(100, 0 , 1); descriptive(x)`
  
+ testing `combinations()`:<br>
  `choose(n = 5, k = 2)` = `r choose(n = 5, k = 2)`<br>
  `combinations(n = 5, k = 2)` = `r combinations(n = 5, k = 2)`

+ testing `binom_prob()`:<br>
  `binom_prob(n = 5, k = 2, prob = 0.5)` = `r binom_prob(n = 5, k = 2, prob = 0.5)`
  `<br>
  more than 3 heads in 5 tosses, with a bais of 35%: `r binom_prob(5, 4, .35) + binom_prob(5, 5, .35)`<br>
  getting 3/6 sixes: `r binom_prob(10, 3, 1/6)`
  
```{r}
squish_0_100 <- function(x){
  if (x < 0){
    return(0)
  } else if (x > 100){
    return(100)
  } else {
    return(x)
  }
}

is_even <- function(x){
  if (is.numeric(x)){
    if (x %% 2 == 0){
      return(TRUE)
    } else {
      return(FALSE)
    }
  } else {
    return(NA)
  }
}

is_odd <- function(x){
  if (is.na(is_even(x))){
    return(NA)
  } else if (is_even(x)){
    return(FALSE)
  } else {
    return(TRUE)
  }
}
```

+ testing `squish`:<br>
  x = -5: `r squish_0_100(-5)`<br>
  x = 51: `r squish_0_100(51)`<br>
  x = 175: `r squish_0_100(175)`
  
+ testing `is_even()`:<br>
  `is_even(10)`: `r is_even(10)`<br>
  `is_even(33)`: `r is_even(33)`<br>
  `is_even('a')`: `r is_even('a')`
  
+ testing `is_odd()`:<br>
  `is_odd(1)`: `r is_odd(1)`<br>
  `is_odd(4)`: `r is_odd(4)`<br>
  `is_odd('a')`: `r is_odd('a')`

```{r}
grade <- function(score){
  if (score > 100 | score < 0){
    stop("score must be a number between 0 and 100")
  } else if (score <= 100 & score >= 90){
    return("A")
  } else if (score < 90 & score >= 80){
    return("B")
  } else if (score < 80 & score >= 70){
    return("C")
  } else if (score < 70 & score >= 60){
    return("D")
  } else {
    return("F")
  }
}
```

+ testing `grade()`:<br>
  `grade(score = 90)`: `r grade(score = 90)`<br>
  `grade(score = 89.9999)`: `r grade(score = 89.9999)`<br>
  `grade(score = 70.0000001)`: `r grade(score = 70.0000001)`<br>
  `grade(score = 50)`: `r grade(score = 50)`<br>
  `grade(score = -5)`: `r #grade(score = -5)`<br>
  `grade(score = 101)`: `r #grade(score = 101)`

```{r}
miles2inches <- function(x = 1){
  x * 63360
}

miles2feet <- function(x = 1){
  x * 5280
}

miles2yards <- function(x = 1){
  x * 1760
}

miles2meters <- function(x = 1){
  x / 0.00062137
}

miles2kms <- function(x = 1){
  x / 0.62137
}

convert <- function(mi, to = "km"){
  switch(to,
         "in" = miles2inches(mi),
         "ft" = miles2feet(mi),
         "yd" = miles2yards(mi),
         "m" = miles2meters(mi),
         "km" = miles2kms(mi),
         NA)
}
```

+ testing convertion functions:<br>
  `miles2inches(2)`: `r miles2inches(2)`<br>
  `miles2feet(2)`: `r miles2feet(2)`<br>
  `miles2yards(2)`: `r miles2yards(2)`<br>
  `miles2meters(2)`: `r miles2meters(2)`<br>
  `miles2kms(2)`: `r miles2kms(2)`
  
+ testing `convert()`:<br>
  `convert(3, "in")`: `r convert(3, "in")`<br>
  `convert(3, "ft")`: `r convert(3, "ft")`<br>
  `convert(3, "yd")`: `r convert(3, "yd")`<br>
  `convert(3, "m")`: `r convert(3, "m")`<br>
  `convert(3, "km")`: `r convert(3, "km")`
