---
title: 'Lab 2: Vectors and other data structures'
subtitle: "Stat 133, Fall 2017"
author: "Sudarshan Srirangapatanam"
date: "September 1, 2017"
output: github_document
---

```{r setup}
require(knitr)
rm(list = ls()) # Cleaning workplace

load("data/nba2017-salary-points.RData") # Load data

check_length <- function(...){
  num <- length(list(...))
  lengths <- 0
  for (i in 1:num){
    lengths[i] <- length(list(...)[[i]])
  }
  for (j in 1:(num-1))
  return("All objects are of Equal length")
}

four <- player[1:4] # First four players
```

## Variable used and descriptions

Data is taken from [`nba2017-salary-points`](https://github.com/Sudarshan-UCB/stat-133-fall2017/blob/master/labs/Lab02/data/nba2017-salary-points.RData)

- `player`: name of the player.<br>
  Class of object: `r class(player)`<br>
  This object is a Vector: `r is.vector(player)`<br>
  Summary (Length): `r summary(player)[1]`
  
- `team`: team name abbreviation.<br>
  Class of object: `r class(team)`<br>
  This object is a Vector: `r is.vector(team)`<br>
  Summary: List of Teams that correspond to the player.
  
- `position`: player position.<br>
  Class of object: `r class(position)`<br>
  This object is a Vector: `r is.vector(position)`<br>
  Summary (Length): `r summary(position)[1]`
  
- `salary`: salary (in dollars).<br>
  Class of object: `r class(salary)`<br>
  This object is a Vector: `r is.vector(salary)`<br>
  Summary (Min, Median, Mean, Max): `r summary(salary)[c(1,3,4,6)]`
  
- `points`: total scored points.<br>
  Class of object: `r class(points)`<br>
  This object is a Vector: `r is.vector(points)`<br>
  Summary (Min, Median, Mean, Max): `r summary(points)[c(1,3,4,6)]`
  
- `points1`: number of free throws, worth 1 point each.<br>
  Class of object: `r class(points1)`<br>
  This object is a Vector: `r is.vector(points1)`<br>
  Summary (Min, Median, Mean, Max): `r summary(points1)[c(1,3,4,6)]`
  
- `points2`: number of 2-point field goals, worth 2 points each.<br>
  Class of object: `r class(points2)`<br>
  This object is a Vector: `r is.vector(points1)`<br>
  Summary (Min, Median, Mean, Max): `r summary(points2)[c(1,3,4,6)]`
  
- `points3`: number of 3-point field goals, worth 3 points each.<br>
  Class of object: `r class(points3)`<br>
  This object is a Vector: `r is.vector(points3)`<br>
  Summary (Min, Median, Mean, Max): `r summary(points3)[c(1,3,4,6)]`
  
**Do all the objects have the same length?**<br>
`r check_length(player, team, position, salary, points, points1, points2, points3)`

**First four player:**<br>
`r four`

**What happens if you specify:**

- an index of zero: `four[0]`?<br>
  `r four[0]`<br>
  We always get a "0" as an answer with the same class as the original vector. Simply put we get a null vector.

- a negative index: `four[-1]`?<br>
  `r four[-1]`<br>
  The value of index associated with the negative sign is removed.

- various negative indices: `four[-c(1,2,3)]`?<br>
  `r four[-c(1,2,3)]`<br>
  All indexes associated with negative is removed.

- an index greater than the length of the vector: `four[5]`?<br>
  `r four[5]`<br>
  We always get a "NA" as an answer.
  
- repeated indices: `four[c(1,2,2,3,3,3)]`?<br>
  `r four[c(1,2,2,3,3,3)]`<br>
  The value of the index is used everytime an index is encountered. Repeated indices results in repeated value.

## Using `seq()` and `rep()`:
```{r seq_rep}
even <- seq(from = 0, to = length(player), by = 2) # all even indexes
odd <- seq(from = 1, to = length(player), by = 2) # all odd indexes
fives <- seq(from = 0, to = length(player), by = 5) # multiples of five
tens <- seq(from = 0, to = length(player), by = 10) # multiples of tens 
even_rev <- rev(seq(from = 0, to = length(player), by = 2)) # all even indexes in reverse order

player_even <- player[even]
salary_odd <- salary[odd]
team_fives <- team[fives]
points_tens <- points[tens]
team_even_rev <- team[even_rev]

```

*Only a few first and last elements are shown*

all the even elements in `player`:<br>
`r c(head(player_even), tail(player_even))`

all the odd elements in `salary`:<br>
`r c(head(salary_odd), tail(salary_odd))`

all multiples of 5 (e.g. 5, 10, 15, etc) of `team`:<br>
`r c(head(team_fives), tail(team_fives))`

elements in positions 10, 20, 30, 40, etc of `points`:<br>
`r c(head(points_tens), tail(points_tens))`

all the even elements in `team` but this time in reverse order:<br>
`r c(head(team_even_rev), tail(team_even_rev))`

## Logical Subsetting
```{r logical}
logical_a <- four[c(TRUE, TRUE, TRUE, TRUE)]
logical_b <- four[c(TRUE, TRUE, FALSE, FALSE)]
logical_c <- four[c(FALSE, FALSE, TRUE, TRUE)]
logical_d <- four[c(TRUE, FALSE, TRUE, FALSE)]
logical_e <- four[c(FALSE, FALSE, FALSE, FALSE)]

recycle_a <- four[TRUE]
recycle_b <- four[c(TRUE, FALSE)]

# Commands to questions after Your Turn
logical_gsw <- player[team == "GSW" & position == "C"]
logical_gsw_lal <- player[team == "GSW" | team == "LAL"]
logical_lal_sg_pg_lal <- player[team == "LAL" & (position == "SG" | position == "PG")]
logical_gsw_lal_sf <- player[position == "SF" & (team == "GSW" & team == "LAL")]
logical_high <- player[salary == max(salary)]
logical_low <- player[salary == min(salary)]
logical_high_pt <- player[points == max(points)]
logical_high_pt_sal <- salary[points == max(points)]
salary_high_c <- max(salary[position == "C"])
logical_team_high_pt <- team[points == max(points)]
logical_high_3pt <- player[points3 == max(points3)]

# Subsetting with character vector
warriors_player <- player[team == "GSW"]

warriors_salary <- salary[team == "GSW"]
warriors_points <- points[team == "GSW"]
names(warriors_salary) <- warriors_player # Assinging names
```

* `four[c(TRUE, TRUE, TRUE, TRUE)]`:<br>
  `r logical_a`

* `four[c(TRUE, TRUE, FALSE, FALSE)]`:<br>
  `r logical_b`
  
* `four[c(FALSE, FALSE, TRUE, TRUE)]`:<br>
  `r logical_c`

* `four[c(TRUE, FALSE, TRUE, FALSE)]`:<br>
  `r logical_d`

* `four[c(FALSE, FALSE, FALSE, FALSE)]`:<br>
  `r logical_e`
  
* `four[TRUE]`:<br>
  `r recycle_a`

* `four[c(TRUE, FALSE)]`:<br>
  `r recycle_b`
  
* players in position Center, of Warriors (GSW):<br>
  `r logical_gsw`

* players of both GSW (warriors) and LAL (lakers):<br>
  `r logical_gsw_lal`

* players in positions Shooting Guard and Point Guards, of Lakers (LAL):<br>
  `r logical_lal_sg_pg_lal`

* subset Small Forwards of GSW and LAL:<br>
  `r logical_gsw_lal_sf`

* name of the player with largest salary:<br>
  `r logical_high`

* name of the player with smallest salary:<br>
  `r logical_low`

* name of the player with largest number of points:<br>
  `r logical_high_pt`
  
* salary of the player with largest number of points:<br>
  `r logical_high_pt_sal`

* largest salary of all Centers:<br>
  `r salary_high_c`

* team of the player with the largest number of points:<br>
  `r logical_team_high_pt`

* name of the player with the largest number of 3-pointers:<br>
  `r logical_high_3pt`

* `warriors_player`:<br>
  `r warriors_player`
  
* `warriors_salary`:<br>
  `r warriors_salary`
  
* `warriors_points`:<br>
  `r warriors_points`
  
* Using Character subsetting:<br>
  `r warriors_salary["Andre Iguodala"]`<br>
  `r warriors_salary[c("Stephen Curry", "Kevin Durant")]`

## Plotting

**Scatter plot of `points` and `salary`:**
```{r scatter}
plot(points, salary)
```

**Looking at the generated plot, can you see any issues?**<br>
Plot is too crowded on one corner but not ont he other corner. Plot doesn't have enough resolution to estimate a pattern at first sight.

**Scatter plot using log transformed vectors:**
```{r scatter_log}
log_points <- log(points)
log_salary <- log(salary)
plot(log_points, log_salary)
#text(log_points, log_salary, labels = abbreviate(player)) # Adding and shortening names
```

## Recycling
```{r recycle}
salary_millions <- salary / (10^6)

#height_cms <- height * 2.54 # This is not given the data
```

## Factors
```{r factor}
position_fac <- as.factor(position)

# Manipulating factors
factor_a <- position_fac[team == "GSW"]
factor_b <- position_fac[salary_millions > 15]
factor_c <- table(position_fac[salary_millions > 15])
factor_d <- prop.table(table(team[position_fac == 'SG']))
```

**Is team a factor:** `r is.factor(team)`

Frequencies of positions:<br>
`r kable(as.matrix(table(position_fac)))`

**manipulating factors:**

* positions of Warriors:<br>
  `r factor_a`

* positions of players with salaries > 15 millions:<br>
  `r factor_b`

* frequencies (counts) of positions with salaries > 15 millions:<br>
  `r kable(as.matrix(factor_c))`

* relative frequencies (proportions) of 'SG' (Shooting Guards) in each team:<br>
  `r kable(as.matrix(factor_d))`
  
## Colored Plot
```{r plot_color}
plot(log_points, log_salary, col = position_fac, xlab = "log of Points", ylab = "log of Salary")
```