---
title: "Lab 1: Basics of R"
subtitle: "Stat 133, Fall 2017"
author: "Sudarshan Srirangapatanam"
date: "September 1, 2017"
output: github_document
---
## Expenses

```{r expenses}
phone <- 80
transportation <- 20
groceries <- 527
gym <- 10
rent <- 1500
other <- 83

total <- phone + transportation + groceries + gym + rent + other # Total Expenses
semester <- total * 5 # Semester Expenses
year <- total * 10 # Year Expenses

expenses <- c("Phone" = phone, "Transportation" = transportation, "Groceries" = groceries, "Gym" = gym, "Rent" = rent, "Other" = other)
sorted_exp <- sort(expenses, decreasing = TRUE) # Sorting expenses
```

Total Expenses = `r total`<br>
Semester Expenses = `r semester`<br>
Year Expenses = `r year`

Expenses Vector:<br>
`r expenses`

**Barplot and plot sorted plus text:**

```{r barplot}
barplot(expenses, names.arg = "")

bar_plot <- barplot(sorted_exp, ylim = c(0, 2000))
plot_text <- bar_plot + text(bar_plot, sorted_exp, labels = sorted_exp, pos = 3)
```

## Pythagoras formula

```{r pythagoras}
a <- 3
b <- 4
c <- sqrt(a^2 + b^2)
```

$$c = \sqrt{a^2 + b^2}$$
a = `r a`, b = `r b`, c = `r c`

## Binomial Formula

```{r binom, echo}
n <- 5 # number of trials
k <- 2 # number of successes
p <- .5 # prob. of a success

## using factorial method
n_k <- factorial(n)/(factorial(k)*factorial(n-k))
coin_fact <- n_k * p^k * p^(n-k)

## using choose method
coin_choose <- choose(n, k) * p^k * p^(n-k)

dice_choose <- choose(10, 3) * (1/6)^3 * (1 - 1/6)^(10-3) # 3 out of 10 six on a dice

## using dbinom method
coin_binom <- dbinom(2, 5, .5) # 2 out of 5 heads
dice_binom <- dbinom(3, 10, 1/6) # 3 out of 10 six on a dice
dice_binom_three <- dbinom(3, 3, 1/6) # 3 out of 3 six on a dice

# To calculate probability with baised coing P(head) = .35
coin_baised <- dbinom(2, 5, .35)

# More than 3 heads with P(head) = .35
coin_sum <- dbinom(4, 5, .35) + dbinom(5, 5, .35) # OR
coin_sum_cdf <- 1 - pbinom(3, 5, .35)

# Installing package
#install.packages("stringr")
#install.packages("RColorBrewer")
#install.packages("XML")

# Calculating 3x^2 + 4x + 8
x <- 2
eqn_two <- 3*x^2 + 4*x + 8

y <- (-3:3)
eqn_series <- 3*y^2 + 4*y + 8
```

$$Pr(k;n,p) = Pr(X = k) = {n \choose k}p^k (1-p)^{n-k}$$
**2 heads out of 5 coin tosses:**

* Using factorial() method:<br>
  $Pr(2, 5, .5)$ = `r coin_fact`
* Using choose() method:<br>
  $Pr(2, 5, .5)$ = `r coin_choose`
* Using dbinom() method:<br>
  $Pr(2, 5, .5)$ = `r coin_binom`

**3 sixes out of 10 dice throws:**

* Using choose() method:<br>
  $Pr(3, 10, \frac{1}{6})$ = `r dice_choose`
* Using dbinom() method:<br>
  $Pr(3, 10, \frac{1}{6})$ = `r dice_binom`
  
**3 sixes out of 3 dice throws:**

* using dbinom() method:<br>
  $Pr(3, 3, \frac{1}{6})$ = `r dice_binom_three`

**Coin tosses of a baised coin Pr(heads) = .35:**

* $Pr(2,5,.35)$ = `r coin_baised`
* Using dbinom() method:<br>
  $Pr(X > 3) = Pr(4, 5, .35) + Pr(5, 5, .35)$ = `r coin_sum`
* Using pbinom() method:<br>
  $Pr(X > 3) = 1 - Pr(X \leq 3)$ = `r coin_sum_cdf`
  
**Calculating $3x^2 + 4x + 8$:**

* When x = 2,<br>
  $3x^2 + 4x + 8$ = `r eqn_two`
* When x is a series from -3 to 3,<br>
  $3x^2 + 4x + 8$ = [`r eqn_series`]