---
title: 'Lab 5: First contact with dplyr and ggplot2'
subtitle: "Stat 133, Fall 2017"
author: "Sudarshan Srirangapatanam"
date: "September 13, 2017"
output: github_document
---

## Data

```{r setup}
require("knitr") #used for kable() function
require("dplyr") #used for data wrangling
require("ggplot2") #used fro plots

dat <- read.csv('data/nba2017-players.csv', stringsAsFactors = FALSE)

lal_players <- dat %>% filter(team == "LAL") %>% select(player)
```

**The data used in this lab is taken from [nba2017-players](https://github.com/Sudarshan-UCB/stat-133-fall2017/blob/master/labs/Lab05/data/nba2017-players.csv)**


### Basic "dplyr" verbs

-   use `slice()` to subset the data by selecting the first 5 rows.<br>
`r kable(slice(dat, 1:5), row.names = FALSE)`

-   use `slice()` to subset the data by selecting rows 10, 15, 20, ..., 50.<br>
`r kable(slice(dat, seq(10, 50, 10)), row.names = FALSE)`

-   use `slice()` to subset the data by selecting the last 5 rows.<br>
`r kable(slice(dat, (length(dat)-5+1):length(dat)), row.names = FALSE)`

-   use `filter()` to subset those players with height less than 70 inches tall.<br>
`r kable(filter(dat, height < 70), row.names = FALSE)`

-   use `filter()` to subset rows of Golden State Warriors ('GSW').\*<br>
`r kable(head(filter(dat, team == "GSW")), row.names = FALSE)`

-   use `filter()` to subset rows of GSW centers ('C').\*<br>
`r kable(head(filter(dat, position == "C")), row.names = FALSE)`

-   use `filter()` and then `select()`, to subset rows of lakers ('LAL'), and then display their names.\*<br>
`r head(dat %>% filter(team == "LAL") %>% select(player))[[1]]`

-   use `filter()` and then `select()`, to display the name and salary, of GSW point guards.\*<br>
`r kable(head(dat %>% filter(team == "GSW" & position == "PG") %>% select(player, salary)), row.names = FALSE)`

-   find how to select the name, age, and team, of players with more than 10 years of experience, making 10 million dollars or less.\*<br>
`r kable(head(dat %>% filter(experience > 10 & salary <= 10^7) %>% select(player, age, team)), row.names = FALSE)`

-   find how to select the name, team, height, and weight, of rookie players, 20 years old, displaying only the first five occurrences (i.e. rows).<br>
`r kable((dat %>% filter(experience == 0 & age == 20) %>% select(player, team, height, weight) %>% slice(1:5)), row.names = FALSE)`

*\* only a small suset is shown using the `head()` function to reduce report length*

### Adding new variables: `mutate()`
```{r mutate_order}
gsw <- filter(dat, team == 'GSW')
gsw <- select(gsw, player, height, weight)
gsw <- slice(gsw, c(4, 8, 10, 14, 15))
gsw

gsw3 <- mutate(gsw, log_height = log(height), log_weight = log(weight))

gsw_mpg <- dat %>% filter(team == "GSW") %>% mutate(min_per_game = minutes / games) %>% select(player, experience, min_per_game) %>% arrange(desc(min_per_game))
```

-   using the data frame `gsw`, add a new variable `product` with the product of `height` and `weight`.<br>
`r kable(mutate(gsw, product = height * weight))`

-   create a new data frame `gsw3`, by adding columns `log_height` and `log_weight` with the log transformations of `height` and `weight`.<br>
`r kable(gsw3)`

-   use the original data frame to `filter()` and `arrange()` those players with height less than 71 inches tall, in increasing order.<br>
`r kable((dat %>% filter(height < 71) %>% arrange(height)), row.names = FALSE)`

-   display the name, team, and salary, of the top-5 highest paid players<br>
`r kable((dat %>% select(player, team, salary) %>% arrange(desc(salary)) %>% slice(1:5)), row.names = FALSE)`

-   display the name, team, and points3, of the top 10 three-point players<br>
`r kable((dat %>% select(player, team, points3) %>% arrange(desc(points3)) %>% slice(1:10)), row.names = FALSE)`

-   create a data frame `gsw_mpg` of GSW players, that contains variables for player name, experience, and `min_per_game` (minutes per game), sorted by `min_per_game` (in descending order).\*<br>
`r kable(head(gsw_mpg), row.names = FALSE)`

*\* only a small suset is shown using the `head()` function to reduce report length*

### Summarizing values with `summarise()`

-   use `summarise()` to get the largest height value.<br>
`r dat %>% summarise(max(height))`

-   use `summarise()` to get the standard deviation of `points3`.<br>
`r dat %>% summarise(sd(points3))`

-   use `summarise()` and `group_by()` to display the median of three-points, by team.<br>
`r kable((dat %>% group_by(team) %>% summarise(median = median(points3))))`

-   display the average triple points by team, in ascending order, of the bottom-5 teams (worst 3pointer teams)<br>
`r kable(dat %>% group_by(team) %>% summarise(average = mean(points3)) %>% arrange(average) %>% slice(1:5))`

-   obtain the mean and standard deviation of `age`, for Power Forwards, with 5 and 10 years (including) years of experience.<br>
`r dat %>% filter(position == "PF" & (experience %in% 5:10)) %>% summarise(mean = mean(age), std_error = sd(age))`

### First contact with `ggplot()`
```{r contact_ggplot}
ggplot(gsw)+
  geom_point(aes(x = height, y = weight))

ggplot(gsw, aes(x = height, y = weight))+
  geom_point()+
  geom_text(aes(label = player))

dat %>% filter(team == "GSW") %>% ggplot(aes(x = height, y = weight))+
  geom_label(aes(label = player))

ggplot(dat, aes(x = salary))+
  geom_density()

ggplot(dat, aes(x = points2))+
  geom_histogram(binwidth = 50)

ggplot(dat, aes(x = position))+
  geom_bar()

dat %>% filter(position == "C") %>% ggplot(aes(x = experience, y = salary))+
  geom_point()+
  geom_smooth(method = "lm")

dat %>% filter(position == "C") %>% ggplot(aes(x = experience, y = salary))+
  geom_point()+
  geom_smooth(method = "loess")
```

### Faceting
```{r faceting}
ggplot(dat, aes(x = experience, y = salary))+
  geom_point()+
  facet_wrap(~ position)

ggplot(dat, aes(x = experience, y = salary))+
  geom_point()+
  facet_wrap(~ team)

ggplot(dat, aes(x = age))+
  geom_density()+
  facet_wrap(~ team)

ggplot(dat, aes(x = height, y = weight))+
  geom_point()+
  facet_wrap(~ position)

ggplot(dat, aes(x = height, y = weight))+
  geom_point()+
  geom_density2d()+
  facet_grid(~ position)
```
