---
title: 'Lab 7: Simple Loops'
subtitle: "Stat 133, Fall 2017"
author: "Sudarshan Srirangapatanam"
output: github_document
---

## Setup

```{r setup}
library("knitr")
library("dplyr")
library("ggplot2")
library("readr")

knitr::opts_chunk$set(fig.path = "../images/")
```

+ `knitr` is used for knitting the document as well as other fucntions to cleanup the output
+ `readr` is used for reading data into R
+ `dplyr` is used for data wrangling
+ `ggplot2` is used to generate any necessary plots

```{r intro}
vec <- c(3, 1, 4)
for (i in 1:length(vec)){
  print(vec[i] * 3)
}

vec2 <- NULL
for (i in 1:length(vec)){
  vec2[i] <- vec[i] * 3
  print(vec2[i])
}
```

```{r summation}
getsum <- function(x, n){
  total <- 0
    for (i in 0:n){
        total <- total + (1 / x^i)
    }
  return(total)
}

vec_getsum <- function(x, a){
  vec <- NULL
  for (i in 1:length(a)){
    vec [i] <- getsum(x, a[i])
  }
  return(vec)
}

vec_getsum(2, 0:5)
```

$$\sum_{k=0}^{n} \frac{1}{2^k} = 1 + \frac{1}{2} + \frac{1}{4} + \frac{1}{8} + \dots + \frac{1}{2^n}$$
`r n <- 0:10; kable(data.frame("n" = n, "sum" = vec_getsum(2, n)))`

$$\sum_{k=0}^{n} \frac{1}{9^k} =1 + \frac{1}{9} + \frac{1}{81} + \dots + \frac{1}{9^n}$$
`r n <- 0:10; kable(data.frame("n" = n, "sum" = vec_getsum(9, n)))`

```{r series}
arith <- function(n = 1, a1 = 3, d = 3){
  a <- NULL
  for (i in 1:n){
    a[i] <- a1 + ((i - 1) * d)
  }
  return(a)
}

geom_sum <- function(n = 1, a1 = 3, r = 2){
  a <- NULL
  for (i in 1:n){
    a[i] <- a1 * (r^(i - 1))
  }
  return(sum(a))
}

vec_geom_sum <- function(a, a1 = 3, r = 2){
  vec <- NULL
  for (i in 1:length(a)){
    vec [i] <- geom_sum(n = a[i], a1, r)
  }
  return(vec)
}
```

`r arith(10, 3, 3)`...<br>
The series doesn't converge.

`r n <- 1:10; kable(data.frame("n" = n, "sum" = vec_geom_sum(a = n)))`

```{r approximation}
sin_x <- function(x, terms = 1){
  sign <- rep(c(1, -1), times = terms / 2)
  if (terms %% 2 == 1){
    sign <- c(sign, 1)
  }
  div <- seq(from = 1, by = 2, length.out = terms)
  
  total <- 0
  for (i in 1:terms){
  total <- total + ((sign[i] * x^(div[i])) / factorial(div[i]))
  }
  return(total)
}
```

using 50 terms:<br>
`r x <- seq(1,2, by = .2) * pi; kable(data.frame("x" = x, "sin_x" = sin_x(x, terms = 50), "sin" = sin(x)))`

using 10 terms:<br>
`r x <- seq(1,2, by = .2) * pi; kable(data.frame("x" = x, "sin_x" = sin_x(x, terms = 10), "sin" = sin(x)))`

using 5 terms:<br>
`r x <- seq(1,2, by = .2) * pi; kable(data.frame("x" = x, "sin_x" = sin_x(x, terms = 5), "sin" = sin(x)))`

```{r matrix_loop}
mod_matrix <- function(A){
    for (i in 1:nrow(A)){
    for (j in 1:ncol(A)){
      if (A[i,j] < 0){
        A[i,j] <- A[i,j]^2
      } else {
        A[i,j] <- sqrt(A[i,j])
      }
    }
    }
  return(A)
}
```

`r kable(mod_matrix(matrix(rnorm(12), nrow = 4, ncol = 3)))`

```{r while_loops}
reduce <- function(x){
  while (x %% 2 == 0){
    x <- x / 2
    print(x)
  }
}
reduce(898128000)
```



```{r avg}
avg_for <- function(x){
  total <- 0
  for (i in 1:length(x)){
    total <- total + x[i]
  }
  return(total / length(x))
}

avg_while <- function(x){
  total <- 0
  i <- 1
  while (i <= length(x)){
    total <- total + x[i]
    i <- i + 1
  }
  return(total / length(x))
}


avg_repeat <- function(x){
  total <- 0
  i <- 1
  repeat {
    total <- total + x[i]
    i <- i + 1
    if (i > length(x)){
      break()
    }
  }
  return(total / length(x))
}
```

$$ \bar{x} = \frac{1}{n} \sum_{i=1}^{n} x_i = \frac{x_1 + x_2 + \dots + x_n}{n} $$

using `for` loop: `r avg_for(1:100)`<br>
using `while` loop: `r avg_while(1:100)`<br>
using `repeat` loop: `r avg_repeat(1:100)`

```{r sd}
sd_for <- function(x){
  total <- 0
  for (i in 1:length(x)){
    total <- total + (x[i] - mean(x))^2
  }
  return(sqrt(total / (length(x) - 1)))
}

sd_while <- function(x){
  total <- 0
  i <- 1
  while (i <= length(x)){
    total <- total + (x[i] - mean(x))^2
    i <- i + 1
  }
  return(sqrt(total / (length(x) - 1)))
}


sd_repeat <- function(x){
  total <- 0
  i <- 1
  repeat {
    total <- total + (x[i] - mean(x))^2
    i <- i + 1
    if (i > length(x)){
      break()
    }
  }
  return(sqrt(total / (length(x) - 1)))
}
```

$$ SD = \sqrt{ \frac{1}{n-1} \sum_{i=1}^{n} (x_i - \bar{x})^2 } $$

using `for` loop: `r sd_for(1:100)`<br>
using `while` loop: `r sd_while(1:100)`<br>
using `repeat` loop: `r sd_repeat(1:100)`

```{r geom}
geom_for <- function(x){
  prod <- 1
  for (i in 1:length(x)){
    prod <- prod * x[i]
  }
  return(prod^(1/length(x)))
}

geom_while <- function(x){
  prod <- 1
  i <- 1
  while (i <= length(x)){
    prod <- prod * x[i]
    i <- i + 1
  }
  return(prod^(1/length(x)))
}


geom_repeat <- function(x){
  prod <- 1
  i <- 1
  repeat {
    prod <- prod * x[i]
    i <- i + 1
    if (i > length(x)){
      break()
    }
  }
  return(prod^(1/length(x)))
}
```

$$ \bar{x} = \left ( \prod_{i=1}^{n} x_i \right )^{1/n} $$

using `for` loop: `r geom_for(1:50)`<br>
using `while` loop: `r geom_while(1:50)`<br>
using `repeat` loop: `r geom_repeat(1:50)`


